---
title: "Summary of Data Processing"
output: 
  html_document: default
params:
  upload: NA
  processed: NA
  C13_ID: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# params list passed to report should have uploaded data, processed data, and the identifier for c13.
# If additional parameters need to be included in the future, simply pass them to report() as a named element of the 'params' argument, then access them in this file using params$<newvaluename>

# get parameters passed to report()
upload <- params$upload
processed <- params$processed
c13 <- params$C13_ID

library(dplyr)
library(kableExtra)
library(readr)

panderOptions('knitr.auto.asis', FALSE)

# tables of calculation options, column names, and display names
# calc_funs <- read.csv("calculation_options.csv", stringsAsFactors = FALSE)
calc_cols <- read.csv("calculation_variables.csv", stringsAsFactors = FALSE)
```

###**Input Data Summary**###

Your data started with **`r nrow(upload$f_data)` samples** and measurements for **`r nrow(upload$e_data)` peaks**.

```{r}
# get data scale: will almost always be 'Raw Abundance'
scale <- switch(attributes(upload)$data_info$data_scale,
                "abundance" = "Raw Abundance",
                "log2" = "log-2 Abundance",
                "log10" = "log-10 Abundance",
                "log" = "log-e Abundance",
                "pres" = "Presence-Absence")

# proportion missing
pct_mis <- mean(upload$e_data[-which(colnames(upload$e_data) == getEDataColName(upload))] == 0)
```

**`r pct_mis*100`**% of the observations in the data file were missing.

Measurements were of the form:  **`r scale`**. 

Metadata columns before processing included the following: 

```{r}
# get all non-null column names that are not edata_cname, fdata_cname, mass_cname
temp <- attributes(upload)$cnames[which(!sapply(attributes(upload)$cnames, is.null))]
temp <- temp[which(!(names(temp) %in% c("edata_cname", "fdata_cname", "mass_cname")))]
```

**`r temp`**

****

The following column selections were made for mass, isotopic notation/symbol , molecule form and/or elemental counts:

```{r}
# display names
selection_names <- c("Mass", "Isotopic Notation/Symbol", "Molecule Formula", "Carbon", "Hydrogen", "Oxygen", "Nitrogen", "Phosphorus")

# actual attribute names
selection_values <- list("mass_cname", "isotopic_cname", "mf_cname", "c_cname", "h_cname", "o_cname", "n_cname", "p_cname")

# make vector of columns that the user selected.  'No Selection' if they didnt make a selection
selection_values <- lapply(selection_values, function(name){
  res <- ifelse(is.null(attributes(processed)$cnames[[name]]), "No Selection", attributes(processed)$cnames[[name]])
  if(name == "isotopic_cname") res <- paste0(res, " : ", c13)
  
  res
  }) %>% unlist()

kable(data.frame("Value" = selection_names, "Column Selected" = selection_values))
```

****

###**Processed Variables**###

```{r}
# find the columns that were added to the original data
proc_cols <- processed$e_meta %>% 
  dplyr::select(which(colnames(processed$e_meta) %in% calc_cols$ColumnName)) %>% 
  colnames() %>% setdiff(temp)

### **Reference html strings**

href1 <- "<a href = 'https://pubs.acs.org/doi/pdf/10.1021/ac010560w'> [Hughey et al., 2001]</a>"
href2 <- "<a href = 'https://onlinelibrary.wiley.com/doi/abs/10.1002/rcm.2386'> [Koch & Dittmar, 2006]</a><a href = 'https://onlinelibrary.wiley.com/doi/full/10.1002/rcm.7433'> | [Erratum]</a>"
href3 <- "<a href = 'https://www.sciencedirect.com/science/article/pii/S0016703711000378'> [LaRowe & Van Cappellen, 2011]</a>"
href4 <- "<a href = 'https://pubs.acs.org/doi/abs/10.1021/ac034415p'> [Kim et al., 2003]</a>"
href5 <- "<a href = 'https://www.sciencedirect.com/science/article/pii/S0038071716306447'> [Bailey et al., 2017]</a>"
href6 <- "<a href = 'https://pubs.acs.org/doi/abs/10.1021/acs.analchem.8b00529'> [Rivas-Ubach et al., 2018]</a>"

# list of equations and references, if applicable
equations <- list("$\\frac{O}{C}$", 
                  "$\\frac{H}{C}$", 
                  "$\\frac{N}{C}$", 
                  "$\\frac{P}{C}$", 
                  "$\\frac{N}{P}$",
                  paste0("$IUPAC mass*(14/14.01565)$", href1), 
                  paste0("Nominal Kendrick Mass - Kendrick Mass", href1), 
                  paste0("$-(\\frac{4C + H - 3N - 2O + 5P - 2S}{C})$", href2),
                  paste0("$60.3 - 28.5*NOSC$", href3), 
                  paste0("$\\frac{1 + C - O - S - 0.5*(N + P + H)}{C - O - S - N - P}$", href2), 
                  paste0("$\\frac{1 + C - 0.5O - S - 0.5*(N + P + H)}{C - 0.5*O - S - N - P}$", href2),
                  paste0("$1 + C - O - S - 0.5*(N + P + H)$", href2), 
                  paste0("$1 + C - O - S - 0.5*(N + P + H) - O$", href2), 
                  "Composition of C,H,N,O,S,P",
                  paste0("Compound Class Set 1", href4),
                  paste0("Compound Class Set 2", href5),
                  paste0("Compound Class Set 3", href6)
                  )
```

The following variables were added in the "Preprocessing" tab:

```{r, results = 'asis', message = FALSE}
# get display names that have a corresponding column in the processed columns
cols_for_display <- calc_cols %>% filter(ColumnName %in% proc_cols) %>% purrr::pluck("DisplayName")

# This is only for the element ratios which occupy the first 5 rows of the calculated columns table.
if(all(calc_cols[1:5,]$DisplayName %in% cols_for_display)){
  pander(paste(cols_for_display[1:5]), collapse = ", ")
  pander(" \n\n ")
  pander(paste(unlist(equations)[1:5], collapse = " "))
  pander(" \n\n ")
  }

# for the rest (6 and onward), if the column corresponding to a particular equation is present in the calculated columns, paste an html string:
dontoutput <-lapply(6:length(equations), function(i){
  if(calc_cols$ColumnName[i] %in% proc_cols){
    # paste the display name inside a fixed width, inline-block div
    cat(paste0("<div style = 'display:inline-block;width:275px'>",
               cols_for_display[which(cols_for_display == calc_cols$DisplayName[i])],
               "</div><div style = 'display:inline-block'>")
        )
    # use pander() to add the equation, notice the previous string started another inline-block div
    pander(equations[[i]])
    
    # end the second div and add two breaks
    cat("</div><br><br>")
  }
})

# old code
# cat("<div style = 'display:inline-block;width:275px'>hi</div><div style = 'display:inline-block'>")
# pander(equations[[1]])
# cat("</div>")
```

****

###**Filters**###

```{r}
# get the filters applied and 
filters <- attributes(processed)$filters %>% names()
filters_names <- sapply(filters, function(x){ switch(x, "massFilt" = "Mass Filter", 
                                                     "moleculeFilt" = "Molecule Filter", 
                                                     "formulaFilt" = "Elemental Formula Filter", 
                                                     paste0("Custom Filter on column: ", gsub("emetaFilt_", "", x)))})

num_filtered <- sapply(attributes(processed)$filters, function(x){
  length(x$filtered)
})

filt_df <- data.frame("Filter" = filters_names, "Peaks Removed" = num_filtered, row.names = NULL, check.names = FALSE)
```

The following filters were applied to the data:

**`r if(length(filters_names) > 0) filters_names else "No filters were applied to the data"`** 

`r if(length(filters_names) > 0) "The following table shows the number of peaks removed by each filter. Filters are applied *in order* from top to bottom, i.e. 'Peaks Removed' for any given filter is the peaks removed after previous filters have already been applied." else NULL`

`r if(length(filters_names) > 0) kable(filt_df, format = "markdown") else NULL`

***

After all preprocessing steps, **`r nrow(processed$f_data)` samples** and measurements for **`r nrow(processed$e_data)` peaks** remained.

<!-- ### **References** -->
<!-- <div id = 'refs'></div> -->
<!-- $(1)$ [Hughey et al., 2001](https://pubs.acs.org/doi/pdf/10.1021/ac010560w) -->

<!-- $(2)$ [Koch \& Dittmar, 2006](https://onlinelibrary.wiley.com/doi/abs/10.1002/rcm.2386) | [Erratum](https://onlinelibrary.wiley.com/doi/full/10.1002/rcm.7433) -->

<!-- $(3)$ [LaRowe \& Van Cappellen, 2011](https://www.sciencedirect.com/science/article/pii/S0016703711000378) -->

<!-- $(4)$ [Kim et al., 2003](https://pubs.acs.org/doi/abs/10.1021/ac034415p) -->

<!-- $(5)$ [Bailey et al., 2017](https://www.sciencedirect.com/science/article/pii/S0038071716306447) -->

<!-- $(6)$ [Rivas-Ubach et al., 2018](https://pubs.acs.org/doi/abs/10.1021/acs.analchem.8b00529) -->


